Spatial Reasoning and Statistical Graphics

### 2014-10-12 Paper/Draft1/TestingVisualAptitude.Rnw

\section{Results}
\newtext{Results are based on an evaluation of \Sexpr{nrow(ans)} undergraduate students at Iowa State University. }
Scoring of all test results was done such that random guessing leads to an expected value of 0; thus for a test consisting of multiple choice questions with $k$ suggested answers with a single correct answer each, the score is calculated as
\begin{eqnarray}\label{eq.scoring}
\# \text{correct answers } - 1/(k-1) \cdot \# \text{wrong answers}.
\end{eqnarray}
This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses. Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests \citep{diamond1973correction, ekstrom1976manual} which parts of this test are drawn from.

Additionally, we have to ensure that the ranges and units of test scores are comparable. Assume $n$ questions with $k$ choices (including one correct answer) each. 
This leads to a theoretical range of $[-n/(k-1), n]$ and, under an additional assumption of random guessing, a variance of

\begin{eqnarray*}
\text{Var}(X_{n, k}) &=& n^2 \text{Var}(X_{n, k}) = \\
&=& n^2 (\underbrace{1/k \cdot 1^2}_{\text{correct answer}}+ \underbrace{(-1/(k-1))^2 \cdot (k-1)/k}_{\text{wrong answer}}) = \\
&=& n^2/(k-1)
\end{eqnarray*}

>>>>

\section{Results}
Results are based on an evaluation of \Sexpr{nrow(ans)} undergraduate students at Iowa State University. 
Scoring of all test results was done such that random guessing leads to an expected value of 0; therefore each question answered correctly contributes to the score by 1, while a wrong answer is scored by $-1/(k-1)$. Thus, for a test consisting of multiple choice questions with $k$ suggested answers with a single correct answer each, the score is calculated as
\begin{eqnarray}\label{eq.scoring}
\# \text{correct answers } - 1/(k-1) \cdot \# \text{wrong answers}.
\end{eqnarray}
This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses. Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests \citep{diamond1973correction, ekstrom1976manual} which parts of this test are drawn from.

Additionally, we have to ensure that the ranges and units of test scores are comparable.  Let $X_{n, k}$ be a participant's score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
This leads to a theoretical range of $X_{n,k}$ of $[-n/(k-1), n]$ and, under an additional assumption of random guessing, a variance of

\begin{eqnarray*}
\text{Var}(X_{n, k}) &=& n^2 \text{Var}(X_{1, k}) = \\
&=& n^2 (\underbrace{1/k \cdot 1^2}_{\text{correct answer}}+ \underbrace{(-1/(k-1))^2 \cdot (k-1)/k}_{\text{wrong answer}}) = \\
&=& n^2/(k-1).
\end{eqnarray*}

\newtext{The above consideration only assumes independence between questions, which is granted. While we only consider a test consisting of questions with the same number of choices, $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure classification score.}
